Appendix 1: Evaluation of the three species in Figure 1

### 2015-10-15 Harris_2015/manuscript-materials/draft.Rmd
clean up directory

@@

@@

@@

**Figure captions**

**Figure 1.**
**A.** A small network of three competing species. The tree (top) tends not to
co-occur with either of the two shrub species, as indicated by the strongly
negative coefficient linking them. The two shrub species also compete with one
another, but more weakly (circled coefficient). **B.** In spite of the
competitive interactions between the two shrub species, their shared tendency to
occur in locations without trees makes their occurrence vectors positively
correlated (circled). **C.** Controlling for the tree species' presence with a
conditional method such as a partial covariance or a Markov network leads to
correct identification of the negative shrub-shrub interaction (circled).

**Figure 2.**
**A.** A small Markov network with two species.  The abiotic environment favors
the occurrence of both species ($\alpha >0$), particularly species 2
($\alpha_2 > \alpha_1$). The  negative $\beta$ coefficient linking these two
species implies that they co-occur less than expected under independence.
**B.** Relative probabilities of all four possible presence-absence combinations
for Species 1 and Species 2. The exponent includes $\alpha_1$ whenever Species 1
is present ($y_1 = 1$), but not when it is absent ($y_1 = 0$).  Similarly, the
exponent includes $\alpha_2$ only when species $2$ is present ($y_2 = 1$), and
$\beta$ only when both are present ($y_1y_2 = 1$). The normalizing constant $Z$,
ensures that the four relative probabilities sum to 1.  In this case, $Z$ is
about 18.5.  **C.** We can find the expected frequencies of all possible
co-occurrence patterns between the two species of interest. **D.** If
$\beta_{12}$ equaled zero (e.g. if the species no longer competed for the same
resources), then the reduction in competition would allow each species to
increase its occurrence rate and the co-occurrence deficit would be eliminated.

**Figure 3.**
**A.** Proportion of variance in interaction coefficients explained by each
method versus number of sampled locations.
**B.** The $Z$-scores produced by the null model ("Pairs") for each pair of
species can be predicted using the correlation between the presence-absence
vectors of those same species and from the number of sites on the landscape.

>>>>

![**Figure 1.** **A.** A small network of three competing species. The tree (top) tends not to
co-occur with either of the two shrub species, as indicated by the strongly
negative coefficient linking them. The two shrub species also compete with one
another, but more weakly (circled coefficient). **B.** In spite of the
competitive interactions between the two shrub species, their shared tendency to
occur in locations without trees makes their occurrence vectors positively
correlated (circled). **C.** Controlling for the tree species' presence with a
conditional method such as a partial covariance or a Markov network leads to
correct identification of the negative shrub-shrub interaction (circled).](figures/Figure_1.pdf)

@@

![**Figure 2.**
**A.** A small Markov network with two species.  The abiotic environment favors
the occurrence of both species ($\alpha >0$), particularly species 2
($\alpha_2 > \alpha_1$). The  negative $\beta$ coefficient linking these two
species implies that they co-occur less than expected under independence.
**B.** Relative probabilities of all four possible presence-absence combinations
for Species 1 and Species 2. The exponent includes $\alpha_1$ whenever Species 1
is present ($y_1 = 1$), but not when it is absent ($y_1 = 0$).  Similarly, the
exponent includes $\alpha_2$ only when species $2$ is present ($y_2 = 1$), and
$\beta$ only when both are present ($y_1y_2 = 1$). The normalizing constant $Z$,
ensures that the four relative probabilities sum to 1.  In this case, $Z$ is
about 18.5.  **C.** We can find the expected frequencies of all possible
co-occurrence patterns between the two species of interest. **D.** If
$\beta_{12}$ equaled zero (e.g. if the species no longer competed for the same
resources), then the reduction in competition would allow each species to
increase its occurrence rate and the co-occurrence deficit would be eliminated.](figures/Figure_2.pdf)

@@

![**Figure 3.**
**A.** Proportion of variance in interaction coefficients explained by each
method versus number of sampled locations.
**B.** The $Z$-scores produced by the null model ("Pairs") for each pair of
species can be predicted using the correlation between the presence-absence
vectors of those same species and from the number of sites on the landscape.](figures/performance.pdf)

@@
